---
title: "problem set#1"
author: "Ali Altahafi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(forcats)
library(dplyr)
library(yarrr)
library(tidyverse)
```
#loading data 
```{r}
data <- read.csv('yarrr_data.csv')
data
```
#numerical variables, it is 11
```{r}
sapply(data, is.numeric)
```
#character variables are 6 

#2.2, the oldest pirate is 46 

```{r}
max(data$age, na.rm=TRUE)
```
#2.3, calculate the mean, the mean is 170.22
```{r}
mean(data$height)
```
#the median is 169.86 or 170 
```{r}
median(data$height)
```
#the standard deviation is 12.38
```{r}
sd(data$height)
```
#2.4, average number owned by pirates is 2.81 or 3
```{r}
mean(data$parrots)
```
#the average number of parrots owned by sex. male = 2.66, female = 3.03, other = 2.28
```{r}
tapply(data$parrot, data$sex, mean)
```
#cerating a frequency table of beard length
```{r}
data %>%
  select(beard.length) %>%
  table() %>%
  prop.table()
```
#generating a cross-tabulation of sword type and eyepatch 
```{r}
crossdata <- table(data$sword.type, data$eyepatch)
print(crossdata)
```
```{r}
set.seed(100)
library(readr)
faketucky <- read_csv("faketucky.csv")
View(faketucky)
```
#changing student_id into a factor variable.  
```{r}
set.seed(100)
Id_factor <- as.factor(faketucky$student_id)
Id_factor
str(Id_factor)
```
#running a t-test but there is an issue with levels
```{r}
set.seed(100)
#na.omit(faketucky$male)
#na.omit(faketucky$gpa)
faketucky <- faketucky %>%
  mutate(gpa = ifelse(gpa == 999, NA, gpa))
faketucky <- faketucky %>%
  mutate(gender = ifelse(faketucky$male == 1, "Males", "Females"))
t.test(faketucky$gpa ~ faketucky$gender)
```
#ran a linear regression model to test for staistical significance. there is a signficant and negative correlation between gpa and being male. 
```{r}
set.seed(100)
#lm(faketucky$gpa ~ faketucky$gender, data = faketucky) %>%
  #summary()
#make sure i ste it up correctly, I'll switch variables
lm(faketucky$male ~ faketucky$gpa, data = faketucky) %>%
  summary()
#scatter.smooth(faketucky$male, faketucky$gpa)
```










